### 2016-05-04 writing/latex/wilf.tex
edit r3

The classes containing the following patterns are Wilf-equivalent
\begin{equation*}
    \pattern{}{1,2}{} \text{ and }
    \pattern{}{2,1}{}
\end{equation*}
This is true since the only avoiders of these patterns are the decreasing
sequence and the increasing sequence respectively, and both of these
avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
every length.

\nextvar[\varminl]
The classes containing the following patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
It is obvious that these two are Wilf-equivalent since the only
permutations that contain these patterns are \(\perm{1,2}\) and
\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
are counted by the Catalan numbers at all lengths except for length
\(2\) where there is precisely \(1\) avoider.
Therefore the generating function is
\begin{equation*}
  \var(x) = C(x)-x^2
\end{equation*}

\subsection{}
\nextvar
Consider the containers of the patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}

For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n\) - \href{https://oeis.org/A001453}{A001453} offset 2}
\end{equation*}

\subsection{}
Now consider the containers of the patterns
\begin{equation*}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Each of these patterns again occurs precisely once in
any containing permutation. However this time when considering
the region corresponding to the unshaded box we need to take into consideration
\LemmaRef{lem:incdec}
and so the empty box can only contain a decreasing subsequence. There
is precisely one decreasing subsequence at every length, and so there
is exactly one container of each pattern at each length. The three
patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
length \(n\) for all \(n \ge 2\).
This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
\end{equation*}

\subsection{}
The containers of the following patterns
have exactly one occurrence.
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions, \ie all elements in the top region must
be to the right of all elements in the bottom region.
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A014137}{A014137} offset 2}
\end{equation*}

\subsection{}
\nextvar
The trivial Wilf-classes containing the following patterns are
Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not coincident or
symmetries of each other
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

@@

\subsection{}
\nextvar
The classes containing the following patterns are Wilf-equivalent
when avoiders are considered in \(\av{\perm{2,3,1}}\)
\begin{align}
    m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1} &\text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1} \label{eq:singleud}\\
    \text{ and }
    m_3 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/1} &\text{ and }
    m_4 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/0} \label{eq:other}
\end{align}

@@

\subsection{}
\nextvar
The classes containing the following patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/1,1/2,
                    2/0,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/1,0/2,
                    1/0,1/2,
                    2/0,2/1,2/2}
\end{equation*}

@@

\subsection{}
\nextvar
The classes containing the following patterns are Wilf-equivalent
\begin{equation*}
    \begin{aligned}
        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2} &\text{ and }
        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2} \\
        \text{ and }
        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2} &\text{ and }
        m_4 = \pattern{}{1,2}{0/1,1/0,1/1,2/0,2/1,2/2}
    \end{aligned}
\end{equation*}

@@

\subsection{}
\nextvar
The patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
\end{equation*}
are Wilf-equivalent.
If \(\scriptvar_1\) is the set of avoiders of
\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
decomposition around the maximum we have
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
\end{equation*}
This is because \(\scriptvar_1\) is the set of permutations
who have their minimum occur after their maximum.

% If \(\mathcal{M}_2\) is the set of avoiders of
% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
% decomposition around the maximum we have
% \begin{equation*}
%     \mathcal{M}_2 = \varepsilon \sqcup
%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
% \end{equation*}
% Alternatively,
The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
the last element is higher than the penultimate element. This can only occur
if the last element is in a single block in the set partition context, In order
to construct a avoider of length \(n\) we can take any avoider of
\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
block.  This ensures that the last block is never a singleton. This means
that these permutations are also counted by \(\var(x) = 1+xC(x)\).

Considering the last pattern, the only way we can construct an
avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
at the start of the permutation. Adding a new leftmost point in any other
position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
Therefore this is also counted by \(\var(x) = 1+xC(x)\).

@@

We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc \tag{\href{https://oeis.org/A000108}{OEIS: A000108} offset 1}
 \end{equation*}
 This is an offset of the Catalan numbers.

\subsection{}
\nextvar
The patterns
\begin{equation*}
    \begin{aligned}
        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1} \text{ and }
        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}\\
        \text{ and }
        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2} \text{ and }
        m_4 = \pattern{}{2,1}{1/0,1/1,2/1,2/2}\\
    \end{aligned}
\end{equation*}
Can be shown to be Wilf-equivalent. First we consider an occurrence of \(m_2\) in a
permutation in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (2.5, 1) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (2.0, 2) rectangle +(0.5,1);
        \draw (2.5,1) -- (2.5,3);
        \draw (1.5,0.5) node {\(b_1\)};
        \draw (2.25,1.5) node {\(\scriptstyle b_2\)};
        \draw (2.75,2.5) node {\(\scriptstyle b_3\)};
        \modpattern[2]{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}
    \end{tikzpicture}
    }
\end{equation*}
We can choose the lowest occurrence of \(m_2\) without loss of generality.
The region corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
as well as \(\perm{2,1,3}\). The regions corresponding to \(b_2\) and \(b_3\) must
now contain avoiders of \(\perm{2,3,1}\), these regions cannot mix in order to avoid
\(\perm{2,3,1}\). Since we already have an occurrence of
\(m_2\) we do not need to care about creating more occurrences so there are no
other conditions on these boxes.

Now looking at an occurrence of \(m_3\) in \(\pi \in \av{\perm{2,3,1}}\)
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1, 1.5) rectangle +(1,0.5);
        \fill[pattern=north east lines, pattern color=black!75] (2, 1) rectangle +(1,0.5);
        \draw (1,1.5) -- (3,1.5);
        \draw (0.5,0.5) node {\(b_1\)};
        \draw (1.5,1.25) node {\(\scriptstyle b_2\)};
        \draw (2.5,1.75) node {\(\scriptstyle b_3\)};
        \modpattern[2]{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}
    \end{tikzpicture}
    }
\end{equation*}
We consider the leftmost occurrence of \(m_3\), the region corresponding
to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\) as well as
\(\perm{2,3,1}\) once more. The regions corresponding to \(b_2\) and \(b_3\)
must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) these regions cannot
mix, as doing so would lead to an occurrence of \(\perm{2,3,1}\).
Therefore both of these sets of containers are enumerated in the same way.

>>>>

The classes containing following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{} \text{ and }
    \pattern{}{2,1}{}
\end{equation*}
This is true since the only avoiders of these patterns are the decreasing
sequence and the increasing sequence respectively, and both of these
avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
every length.

\nextvar[\varminl]
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
It is obvious that these two are Wilf-equivalent since the only
permutations that contain these patterns are \(\perm{1,2}\) and
\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
are counted by the Catalan numbers at all lengths except for length
\(2\) where there is precisely \(1\) avoider.
Therefore the generating function is
\begin{equation*}
  \var(x) = C(x)-x^2
\end{equation*}

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}
Consider containers of these patterns.
For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the region corresponding to the unshaded box in each case. Each
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n\) - \href{https://oeis.org/A001453}{A001453} offset 2}
\end{equation*}

\subsection{}
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Consider containers of these patterns.
Each of these patterns again occurs precisely once in
any containing permutation. However this time when considering
the region corresponding to the unshaded box we need to take into consideration
\LemmaRef{lem:incdec}
and so the empty box can only contain a decreasing subsequence. There
is precisely one decreasing subsequence at every length, and so there
is exactly one container of each pattern at each length. The three
patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
length \(n\) for all \(n \ge 2\).
This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
\end{equation*}

\subsection{}
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
The containers of the patterns
have exactly one occurrence.
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions, \ie all elements in the top region must
be to the right of all elements in the bottom region.
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A014137}{A014137} offset 2}
\end{equation*}

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{align}
    m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1} &\text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1} \label{eq:singleud}\\
    \text{ and }
    m_3 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/1} &\text{ and }
    m_4 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/0} \label{eq:other}
\end{align}

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/1,1/2,
                    2/0,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/1,0/2,
                    1/0,1/2,
                    2/0,2/1,2/2}
\end{equation*}

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \begin{aligned}
        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2} &\text{ and }
        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2} \\
        \text{ and }
        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2} &\text{ and }
        m_4 = \pattern{}{1,2}{0/1,1/0,1/1,2/0,2/1,2/2}
    \end{aligned}
\end{equation*}

@@

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
\end{equation*}

If \(\scriptvar_1\) is the set of avoiders of
\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
decomposition around the maximum we have
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
\end{equation*}
The lower-left section is empty because the minimum must occur after the maximum.
These are counted by \(\var(x) = 1+xC(x)\)
% This is because \(\scriptvar_1\) is the set of permutations
% who have their minimum occur after their maximum.

% If \(\mathcal{M}_2\) is the set of avoiders of
% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
% decomposition around the maximum we have
% \begin{equation*}
%     \mathcal{M}_2 = \varepsilon \sqcup
%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
% \end{equation*}
% Alternatively,
The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
the last element is higher than the penultimate element. This can only occur
if the last element is in a single block in the set partition context, In order
to construct an avoider of length \(n\) we can take any avoider of
\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
block.  This ensures that the last block is never a singleton. This means
that these permutations are also counted by \(\var(x)\).

Considering the last pattern, the only way we can construct an
avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
at the start of the permutation. Adding a new leftmost point with any other
value would either create an occurrence of \(\perm{2,3,1}\) or the mesh pattern.
Therefore these permutations are also counted by \(\var(x) = 1+xC(x)\).

@@

So clearly \(\var_4(x) = \var_3(x)\).
We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc \tag{\href{https://oeis.org/A000108}{OEIS: A000108} offset 1}
 \end{equation*}
 This is an offset of the Catalan numbers.

\subsection{}
\nextvar
The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    \begin{aligned}
        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1} \text{ and }
        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}\\
        \text{ and }
        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2} \text{ and }
        m_4 = \pattern{}{2,1}{1/0,1/1,2/1,2/2}\\
    \end{aligned}
\end{equation*}
 First we consider an occurrence of \(m_2\) in a
permutation in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (2.5, 1) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (2.0, 2) rectangle +(0.5,1);
        \draw (2.5,1) -- (2.5,3);
        \draw (1.5,0.5) node {\(b_1\)};
        \draw (2.25,1.5) node {\(\scriptstyle b_2\)};
        \draw (2.75,2.5) node {\(\scriptstyle b_3\)};
        \modpattern[2]{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}
    \end{tikzpicture}
    }
\end{equation*}
We can choose the lowest occurrence of \(m_2\) in the sense that the region
corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
as well as \(\perm{2,3,1}\). The regions corresponding to \(b_2\) and \(b_3\) must
now contain avoiders of \(\perm{2,3,1}\), all points in the region corresponding
to \(b_2\) must be to the right of those in the region corresponding to \(b_3\)
\(\perm{2,3,1}\). Since we already have an occurrence of
\(m_2\) we do not need to care about creating more occurrences so there are no
other conditions on these boxes.

Now looking at an occurrence of \(m_3\) in \(\pi \in \av{\perm{2,3,1}}\)
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1, 1.5) rectangle +(1,0.5);
        \fill[pattern=north east lines, pattern color=black!75] (2, 1) rectangle +(1,0.5);
        \draw (1,1.5) -- (3,1.5);
        \draw (0.5,0.5) node {\(b_1\)};
        \draw (1.5,1.25) node {\(\scriptstyle b_2\)};
        \draw (2.5,1.75) node {\(\scriptstyle b_3\)};
        \modpattern[2]{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}
    \end{tikzpicture}
    }
\end{equation*}
We consider the leftmost occurrence of \(m_3\) in the sense that the region
corresponding
to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\) as well as
\(\perm{2,3,1}\) once more. The regions corresponding to \(b_2\) and \(b_3\)
must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) the points in the
region corresponding to the box containing \(b_2\) must be lower in value than
all of those in the region corresponding to the box containing \(b_3\), as doing
so would lead to an occurrence of \(\perm{2,3,1}\).
Therefore both of these sets of containers are enumerated in the same way.

Now we find a structural decomposition for an avoider of \(m_2\).
Decomposing around the maximum we see the set of avoiders of \(m_2\)
have the form
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle\scriptvar_2\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}
We can decompose an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\) around the leftmost point
in a similar manner:
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle\scriptvar_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}
These two decompositions tell us that these two patterns are Wilf-equivalent
and have generating function
\begin{equation}
    \var_1 = 1 + x(\var_1(x)-1)C(x) +x\varfine(x)
    \label{eqn:ngen}
\end{equation}
Now consider an avoider of \(m_4\) decomposed around the maximum\nextvar %subvar
\begin{equation*}
    \mathcal{\lastvar}_4 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptvar}
\end{equation*}
Here \(\scriptvar\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
\(p = \textpattern{}{1}{0/0,0/1,1/1}\), since if the subsequence in this box
were to start with the maximum then this point and the maximum would create
an occurrence of \(m_4\).
Now consider decomposition of a permutation in \(\scriptvar\) around its
leftmost point.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{\lastvar}_4\setminus\varepsilon}
\end{equation*}
This gives us the generating function
\begin{equation*}
    \var(x) = 1 + xC(x)(\lastvar_4(x)-1)
\end{equation*}
Now we get the following for \(\lastvar_4\)
\begin{equation}
    \lastvar_4(x) = 1 +xC(x)(xC(x)(\lastvar_4(x)-1)+1)
    \label{eqn:ngen2}
\end{equation}
